---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RClimacell

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
[![GitHub commit](https://img.shields.io/github/last-commit/nikdata/RClimacell)](https://github.com/nikdata/RClimacell/commit/main)
[![R-CMD-check](https://github.com/nikdata/RClimacell/workflows/R-CMD-check/badge.svg)](https://github.com/nikdata/RClimacell/actions)
<!-- badges: end -->

The {RClimacell} package is an **unofficial** R package that enables basic interaction with [Climacell's](https://www.climacell.co) API using the [Timeline Interface](https://docs.climacell.co/reference/timeline-overview). The functions within this package are tested against some of the [CORE data layers](https://docs.climacell.co/reference/data-layers-core). 

Please note that using the functions within this package **require** a valid API key.

More information about the Climacell API can be found on their [docs](https://docs.climacell.co/reference/api-overview) page.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nikdata/RClimacell")
```

## Usage

### Temperature
```{r temperature_example}
library(RClimacell)
climacell_temperature(api_key = Sys.getenv("CLIMACELL_API"),
                      lat = 41.878685,
                      long = -87.636011,
                      timestep = '1d',
                      start_time = Sys.time(),
                      end_time = Sys.time() + lubridate::days(3))

```

### Wind
```{r wind_example}
library(RClimacell)
climacell_wind(api_key = Sys.getenv("CLIMACELL_API"),
               lat = 41.878685,
               long = -87.636011,
               timestep = '1d',
               start_time = Sys.time(),
               end_time = Sys.time() + lubridate::days(3))
```

## Precipitation
```{r}
library(RClimacell)
climacell_precip(api_key = Sys.getenv("CLIMACELL_API"),
                 lat = 41.878685,
                 long = -87.636011,
                 timestep = '1d',
                 start_time = Sys.time(),
                 end_time = Sys.time() + lubridate::days(3))
```


See the [vignette](https://nikdata.github.io/RClimacell/) for more information.
