---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RClimacell

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![License: MIT](https://img.shields.io/badge/License-MIT-blue.svg)](https://opensource.org/licenses/MIT)
[![GitHub commit](https://img.shields.io/github/last-commit/nikdata/RClimacell)](https://github.com/nikdata/RClimacell/commit/main)
[![R-CMD-check](https://github.com/nikdata/RClimacell/workflows/R-CMD-check/badge.svg)](https://github.com/nikdata/RClimacell/actions)
<!-- badges: end -->

The {RClimacell} package is an unofficial R package that enables basic interaction with [Climacell's](https://www.climacell.co) API. The functions within this package are tested against the [CORE data layers](https://docs.climacell.co/reference/data-layers-core). Using the functions within this package require a valid API key.

**WIP**

This package is still under development. Please use with caution. Functions have not been fully tested.

**THIS IS NOT AN OFFICIAL CLIMACELL PACKAGE!**

More information about the Climacell API can be found on their [docs](https://docs.climacell.co/reference/api-overview) page.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nikdata/RClimacell")
```

## Usage

### Temperature
```{r temperature_example}
library(RClimacell)
climacell_temperature(api_key = Sys.getenv("CLIMACELL_API"),
                      lat = 41.878685,
                      long = -87.636011,
                      timestep = '1d',
                      start_time = Sys.time(),
                      end_time = Sys.time() + lubridate::days(7))

```

### Wind
```{r wind_example}
library(RClimacell)
climacell_wind(api_key = Sys.getenv("CLIMACELL_API"),
                      lat = 41.878685,
                      long = -87.636011,
                      timestep = '1d',
                      start_time = Sys.time(),
                      end_time = Sys.time() + lubridate::days(7))
```


## CAUTION

* The free API has a limit of 1000 calls per day for CORE data layers
* The free API has a limit of 100 calls per day for PREMIUM data layers (out of scope for this package at this time)


## Limitations

* Does not allow for PUT requests
* I don't have the funds to get a more premium API account enabling for richer weather data. Therefore, I am limiting this package (for now) to the CORE data layers and focusing on GET.
* PUT is only needed if you want to modify the settings on your Climacell API account from what I can gather.
